**Covid-19 vaccination in the province of Ontario: A geographical and socio-economical analysis**

### 2023-03-02 manuscript/01-background.qmd
Abstract, discussion, conclusion

In this study we examined self-reported COVID-19 vaccination status within the province of Ontario using a combination of socio-economic factors(such as ethnic origin, age, and income) and geographical analysis (at the level of the Health Regions)  in order to determine intra-provincial disparities in vaccine uptake and address the ongoing need of socio-economic information that can provide a rationale for the disparities in vaccination observed within some racial groups [@cnat2022b].

>>>>

In this study, we examined self-reported COVID-19 vaccination status within the province of Ontario using a combination of socio-economic factors(such as ethnic origin, age, and income) and geographical analysis (at the level of the Health Regions) in order to determine intra-provincial differences in vaccine uptake and address the ongoing need of socio-economic information that can provide a rationale for the disparities in vaccination observed within some racial groups [@cnat2022b].

### 2023-03-02 manuscript/02-methods.qmd
Abstract, discussion, conclusion

For each survey participant certain data was automatically captured, including the nearest municipality (city). This resulted in a total of 578 different municipalities within the clean dataset. Due to our interest in analyzing the differences between Health Regions, we assigned the city of each entry to its correspondent Health Region following a multi-step process. Briefly, we used Local Health Integrated Networks (LHINs) to assign a Health Region to each entry in the survey. LHINs were the geographical divisions for health used by  Ontario before the adoption of the Health Regions; because of the lack of a publicly available list of all municipalities within each Health Region, we used a dataset of long-term care homes and LHINs to match each city to LHIN, followed by matching each LHIN to a Health Region following the information provided on the Ontario Health Website, where the list of LHINs and corresponding Health Regions is available. In the case of municipalities that did not appear in the long-term dataset, we manually searched each city in the LHINs websites in order to provide geographical information. The raw dataset, clean dataset, details of the data cleaning process, and the addition of Health Region and LHIN information are described in detail in the github repository: https://github.com/aimundo/Fields_COVID-19/.

@@

Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a Covid-19 vaccine, $\beta_0$ indicates the population intercept, and $\beta_1...\beta_6$ indicate the coefficients for each term. The the model was fitted using the function `svyglm` from the `survey` R package in order to incorporate the correction in sampling probability obtained from raking.

>>>>

For each survey participant certain data was automatically captured, including the nearest municipality (city). This resulted in a total of 578 different municipalities within the clean dataset. Due to our interest in analyzing the differences between Health Regions, we assigned the city of each entry to its correspondent Health Region following a multi-step process. Briefly, we used Local Health Integrated Networks (LHINs) to assign a Health Region to each entry in the survey. LHINs were the geographical divisions for health used by  Ontario before the adoption of the Health Regions; because of the lack of a publicly available list of all municipalities within each Health Region, we used a dataset of long-term care homes and LHINs to match each city to LHIN, followed by matching each LHIN to a Health Region following the information provided on the Ontario Health Website, where the list of LHINs and corresponding Health Regions is available. In the case of municipalities that did not appear in the long-term dataset, we manually searched each city in the LHINs websites in order to provide geographical information. The raw dataset, clean dataset, details of the data cleaning process, and the addition of Health Region and LHIN information are described in detail in the GitHub repository for this paper, which can be found at [https://github.com/aimundo/Fields_COVID-19/](https://github.com/aimundo/Fields_COVID-19/).

@@

Where $p\textrm{(vac)}$ indicates the probability of having received the first dose of a COVID-19 vaccine, $\beta_0$ indicates the population intercept, and $\beta_1...\beta_6$ indicate the coefficients for each term. The the model was fitted using the function `svyglm` from the `survey` R package in order to incorporate the correction in sampling probability obtained from raking.

### 2023-03-02 manuscript/03-results.qmd
Abstract, discussion, conclusion

@tbl-descriptive-stats shows the descriptive statistics (uncorrected) from the Fields Covid-19 survey data for vaccination status and each of the covariates analyzed. The total number of entries from the survey in the dataset after cleaning was 3,551. Overall, 26.9\% of survey respondents (958) reported not having received the first dose of the vaccine, whereas 73.1\% (2,593) reported having received it. Within each socio-economic factor, respondents who reported living in a household with an income under CAD 25,000 represented  37\% of the total number of entries, those within the CAD 25,000-59,999 income bracket represented 25% of the total sample, and those with an income above CAD 60,000 represented 38 \% of the sample; across all income brackets, the percentage of individuals that reported having received a first dose of the vaccine was consistent, above 69\%.

@@

```{r,summary-table}
#| label: tbl-descriptive-stats
#| tbl-cap: "Descriptive Statistics of the Fields Covid-19 Survey (by Vaccination Status)"
#| message: false
#| echo: false

@@

data_summary<-clean_data %>%
  tbl_summary(percent="row",
              by= first_dose,
              include=c(income_ord,
                        Age_group_ord,
                        Health_Region,
                        Race)
  ) %>% bold_labels()

@@

@tbl-model shows the results of the logistic regression on vaccination status using socio-economic factors (age group, income, race), geographical areas (Health Regions) and the interactions between income and race and Health Region and race. The reference groups were set as follows: 16 to 34 years (age group), White Caucasian (Race), Toronto (Health Region), CAD 60,000 and over (Income). There were no statistically significant differences in vaccination rates within the age groups from the survey, but significant odds ratios were estimated for other covariates. Within household income brackets, individuals with an income under CAD 25,000 or between CAD 25,000-59,999 had significantly lower odds of vaccination than those with an income above CAD 60,000 (ORs=0.37 and 0.59, respectively). Within Race/Ethnicity, individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.31, 0.32, 0.28, respectively); additionally, those individuals in the Other Race/Ethnicity group (a group that included Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere) had even lower odds of vaccination than the other minority groups, with an OR of 0.22. Regarding Health Regions, individuals that reported living in the West Health Region (which comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent) had significantly higher odds of vaccination than those in the Health Region of Toronto (OR=1.55).

Moreover, statistically-significant odd ratios were determined in the case of the interaction of income and race; specifically, for individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.05), Black (OR=3.19), Latin American (OR=2.80), or that belonged to other minority groups  (OR=4.59). Within the CAD 25,000-59,999 income bracket, individuals who identified as belonging to other racial minority groups had significantly higher odds of vaccination (OR=6.93).

For the interaction of Health Region and race, significant odds of vaccination were identified for Black individuals in the Central Health Region, which comprises the region of York, counties of Dufferin and Simcoe and the district of Muskoka (OR=0.44), and in individuals that identified as part of other racial minorities or South Asian that lived in the West Health Region (ORs=0.41).

@@

# create table with the model output
m1 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group")) %>%
    bold_labels()%>%
  as_kable_extra(booktabs = TRUE, longtable=T, linesep="") %>%
  # reduce font size to make table fit. 
  # you may also use the `latex_options = "scale_down" argument here.
  kableExtra::kable_styling(latex_options = "repeat_header")
```

>>>>

@tbl-descriptive-stats shows the descriptive statistics (uncorrected) from the Fields COVID-19 survey data for vaccination status and each of the covariates analyzed. The total number of entries from the survey in the dataset after cleaning was 3,551. Overall, 26.9\% of survey respondents (958) reported not having received the first dose of the vaccine, whereas 73.1\% (2,593) reported having received it. Within each socio-economic factor, respondents who reported living in a household with an income under CAD 25,000 represented  37\% of the total number of entries, those within the CAD 25,000-59,999 income bracket represented 25% of the total sample, and those with an income above CAD 60,000 represented 38 \% of the sample; across all income brackets, the percentage of individuals that reported having received a first dose of the vaccine was consistent, above 69\%.

@@

```{r,summary-table}
#| label: tbl-descriptive-stats
#| tbl-cap: "Descriptive Statistics of the Fields COVID-19 Survey (by Vaccination Status)"
#| message: false
#| echo: false

@@

data_summary<-clean_data %>%
  tbl_summary(percent="row",
              by= first_dose,
              include=c(income_ord,
                        Age_group_ord,
                        Health_Region,
                        Race),
              label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group",
                 income_ord="Income")
  ) %>% bold_labels()

@@

@tbl-model shows the results of the logistic regression on vaccination status using socio-economic factors (age group, income, race), geographical areas (Health Regions) and the interactions between income and race and Health Region and race. The reference groups were set as follows: 16 to 34 years (age group), White Caucasian (Race), Toronto (Health Region), CAD 60,000 and over (Income). There were no statistically significant differences in vaccination rates within the age groups from the survey, but significant odds ratios were estimated for other covariates. Within household income brackets, individuals with an income under CAD 25,000 or between CAD 25,000-59,999 had significantly lower odds of vaccination than those with an income above CAD 60,000 (ORs=0.37 and 0.59, respectively). Within Race/Ethnicity, individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than those in the White/Caucasian group (ORs=0.31, 0.32, 0.28, and _p_=0.004,<0.001 and 0.004, respectively); additionally, those individuals in the Other Race/Ethnicity group (a group that included Southeast Asian, Filipino, West Asian, and Minorities Not Identified Elsewhere) had even lower odds of vaccination than the other minority groups (OR=0.22, _p_<0.001). Regarding Health Regions, individuals that reported living in the West Health Region (which comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex, and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent) had significantly higher odds of vaccination than those in the Health Region of Toronto (OR=1.55, _p_=0.029).

Moreover, statistically-significant odd ratios were determined in the case of the interaction of income and race; specifically, for individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.05, _p_=0.013), Black (OR=3.19, _p_=0.004), Latin American (OR=2.80, _p_=0.041), or that belonged to other minority groups  (OR=4.59, _p_<0.001). Within the CAD 25,000-59,999 income bracket, individuals who identified as belonging to other racial minority groups had significantly higher odds of vaccination (OR=6.93, _p_<0.001).

For the interaction of Health Region and race, significant odds of vaccination were identified for Black individuals in the Central Health Region, which comprises the region of York, counties of Dufferin and Simcoe and the district of Muskoka (OR=0.44, _p_=0.046), and in individuals that identified as part of other racial minorities or South Asian that lived in the West Health Region (ORs=0.41, _p_=0.032 and _p_=0.037, respectively).

@@

# create table with the model output
m1 %>%
  tbl_regression(exponentiate = TRUE,
                 label=list(Health_Region ~ "Health Region",
                 Age_group_ord ~ "Age Group",
                 income_ord="Income")) %>%
    bold_labels()%>%
  as_kable_extra(booktabs = TRUE, longtable=T, linesep="") %>%
  # reduce font size to make table fit. 
  # you may also use the `latex_options = "scale_down" argument here.
  kableExtra::kable_styling(latex_options = "repeat_header")
```

### 2023-03-02 manuscript/04-discussion.qmd
Abstract, discussion, conclusion

The rapid development of COVID-19 vaccines has been considered as a major achievement of modern medicine [@davis2022], and their availability towards the end of 2020 in certain countries made some believe that they would be a determinant factor in a rapid ending of the pandemic [@thelancet2021]. However, despite previous successful vaccination campaigns that were crucial to control diseases such as smallpox and polio [@kayser2021], vaccination efforts in the case of Covid-19 have faced multiple challenges that have complicated the achievement of global immunity.

Among the different challenges faced by Covid-19 vaccination efforts are the development of new variants due to inadequate public health measures [@li2021] and inequity in vaccine access between low and high income countries [@tamey2022]. However, it is also well established that even in the case of high income countries that have had ample access to vaccines since 2020, such as the US, the UK, and Canada, there have been challenges in vaccination efforts due to differences in vaccine uptake among different segments of the population. More specifically, lower vaccine uptake has been associated with socio-economic factors such as race (i.e., identifying as Black, Asian, Indigenous) and household income (typically within lower income brackets) [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021]. Reasons given for this association have included medical mistrust due to systemic medical racism, mistrust in vaccines, and the influence of conspiracy theories [@stoler2021;@willis2021;@bogart2021;@mosby2021;@freeman2020].

In addition, vaccine uptake is influenced by geography as shown by different studies that have identified intra-regional differences in vaccine uptake [@pallathadka2022;@huang2022;@mollalo2021]. However, in the case of Canada, studies that have analyzed spatial differences in vaccination have been focused in country-wide or province-wide estimates [@lavoie2022;@guay2022]. Therefore, we explored spatial and socio-economic determinants of vaccination status in the province of Ontario. This province is of particular interest as it has seen recently major structural health changes with the dissolution of the Local Health Integrated Network (LHIN) system and the incorporation of regions covered by LHINs into larger Health Regions [@dong2022]. Because the idea behind the change aimed to reduce the inequalities in healthcare that were identified under the LHIN model [@tsasis2012], examining differences in vaccination between the Health Regions can provide decision-makers with insight regarding intra-provincial health disparities that may need to be addressed in future vaccination or public health campaigns.

Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of the surveyed individuals reported having received the first dose of the Covid-19 vaccine (@tbl-descriptive-stats), and that there were no significant differences in vaccination odds among the age groups considered in the survey. This result is consistent with overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022], with vaccination uptake rates across different age groups presented in other studies [@macdonald2021;@guay2022], and with the vaccination information provided by Public Health Ontario,  which shows that for the period where the Fields survey ran (Sept 30, 2021-Jan 17, 2022) there was a minimum of  80% of first dose vaccination coverage among all the age groups considered in the survey [@ontario-covid].

@@

This result is likely due to the fact that individuals that belong racial minority groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include occupations such as grocery store workers, gas station workers, warehouse and distribution workers, and manufacturing workers, all being occupations for which a low income is to be expected.  In the case of Ontario, essential workers had priority for Covid-19 vaccination [@mishra2021], which would explain the higher odds of vaccination for these individuals in certain income brackets, in contrast to the lower odds of vaccination for the same type of individuals in higher income brackets. In other words, it is possible that the type of occupation played an important role in the odds of vaccination in these racial minority groups.

Additionally, significant higher vaccination odds were identified in the West Health Region when compared to the Health Region of Toronto (@tbl-model). The West Health Region comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent. In this case, overall higher odds of vaccination can be explained by the fact that in the survey, about 47% of the entries for this Health Region corresponded to White/Caucasian individuals, who reported an overall 83% vaccination rate (Supplementary Table A-6). However, the interaction effect of Health Region and race was also significant in the case of individuals identifying as South Asian or other minorities not included in the survey [@tbl-model]. In this case, the results of the interaction term in the model indicate that the odds of vaccination for those within the South Asian and Other minority groups in the West Region decreased when compared to the other Health Regions (Supplementary Figure A-4).

The West Health Region, according to Ontario Health, has 13.2% of its population identifying as visible minority, and 2.5% of it identifying as Indigenous [@ontariohealth]. In the case of this analysis, it is possible that these significant interactions are due to the fact that 50% of the answers from this region came from the former LHINs of Hamilton Niagara Haldimand Brant, and Erie St. Clair, which were previously reported to have more than 20% of their population in the lowest income quintile [@buajitti2018] (Supplementary Table A-7) while about 55% of the answers from areas previously under the aforementioned LHINs also corresponded to individuals that identified in the minority groups. This would mean that although all minority groups had lower odds of vaccination in the rest of the Health Regions, the combination of higher representation of individuals from low-income areas that identified as minorities resulted in the overall significant lower odds of vaccination.

Although there was no significant difference between vaccination odds among the Health Regions, there was a significant lower vaccination odds in the Central Health Region in individuals with a household income between CAD 15,000-24,999 (OR=0.4, @tbl-model). The Central Health Region comprises the former Central, Central West, Mississauga Halton, and North Simcoe Muskoka LHINs with 48.3% of immigrant population and 47% of the population identifying as visible minority, the highest percentage among all Health Regions [@ontariohealth]; moreover, the fact that the Central West LHIN, now within the Central Health Region, had the highest proportion of low-income families in 2018 (12.9\% of males and 16.7\% of females among the lowest 20% of Canadian family incomes [@buajitti2018])

- Influence of geography

- ontario and Health Regions: find information about poverty and racial groups.

- Results

- Limitations

>>>>

The rapid development of COVID-19 vaccines has been considered as a major achievement of modern medicine [@davis2022]. Vaccine availability towards the end of 2020 in certain countries made some believe that they would be a determinant factor in a rapid ending of the pandemic [@thelancet2021]. However, despite previous successful vaccination campaigns that were crucial to control diseases such as smallpox and polio [@kayser2021], vaccination efforts in the case of COVID-19 have faced multiple challenges that have complicated the achievement of global immunity.

Among the different challenges faced by COVID-19 vaccination efforts are the development of new variants due to inadequate public health measures [@li2021] and inequity in vaccine access between low and high income countries [@tamey2022]. However, it is also well established that even in the case of high income countries that have had ample access to vaccines since 2020, such as the US, the UK, and Canada, there have been challenges in vaccination efforts due to differences in vaccine uptake among different segments of the population. More specifically, lower vaccine uptake has been associated with socio-economic factors such as race (i.e., identifying as Black, Asian, Indigenous) and household income (typically within lower income brackets) [@willis2021;@skirrow2022;@stoler2021;@khubchandani2021]. Reasons given for this association have included medical mistrust due to systemic medical racism, mistrust in vaccines, and the influence of conspiracy theories [@stoler2021;@willis2021;@bogart2021;@mosby2021;@freeman2020].

In addition, vaccine uptake is influenced by geography, as shown by different studies that have identified intra-regional differences in vaccine uptake [@pallathadka2022;@huang2022;@mollalo2021]. However, in the case of Canada, studies that have analyzed spatial differences in vaccination have been focused in country-wide or province-wide estimates [@lavoie2022;@guay2022]. Therefore, we explored spatial and socio-economic determinants of vaccination status in the province of Ontario. This province is of particular interest as it has seen recently major structural health changes with the dissolution of the Local Health Integrated Network (LHIN) system and the incorporation of regions covered by LHINs into larger Health Regions [@dong2022]. Because the idea behind the change aimed to reduce the inequalities in healthcare that were identified under the LHIN model [@tsasis2012], examining differences in vaccination between the Health Regions can provide decision-makers with insight regarding intra-provincial health disparities that may need to be addressed in future vaccination or public health campaigns.

Our results indicate that across the most densely populated Health Regions of Ontario, almost three quarters of the surveyed individuals reported having received the first dose of the COVID-19 vaccine (@tbl-descriptive-stats), and that there were no significant differences in vaccination odds among the age groups considered in the survey. This result is consistent with overall vaccination rates reported for Canada, which have been relatively higher when compared to other high income countries [@dube2022], with vaccination uptake rates across different age groups presented in other studies [@macdonald2021;@guay2022], and with the vaccination information provided by Public Health Ontario,  which shows that for the period where the Fields survey ran (Sept 30, 2021-Jan 17, 2022) there was a minimum of  80% of first dose vaccination coverage among all the age groups considered in the survey [@ontario-covid].

@@

This result is likely due to the fact that individuals that belong racial minority groups tend to perform occupations that have been deemed as "essential" in the context of the pandemic [@hawkins2020;@ct2021], which include occupations such as grocery store workers, gas station workers, warehouse and distribution workers, and manufacturing workers, all being occupations for which an income within the significant brackets is to be expected.  In the case of Ontario, essential workers had priority for COVID-19 vaccination [@mishra2021], which would explain the higher odds of vaccination for these individuals in certain income brackets, in contrast to the lower odds of vaccination for the same type of individuals with higher household income. In other words, it is possible that the type of occupation played an important role in increasing the odds of vaccination in these racial minority groups.

Additionally, significant higher vaccination odds were identified in the West Health Region when compared to the Health Region of Toronto (@tbl-model). The West Health Region comprises the regions of Waterloo and Niagara, the counties of Wellington, Essex and Lambton, and the cities of Hamilton, Haldimand, Brant, and Chatham-Kent. In this case, a possible rationale for the results is the fact that in the survey, about 47% of the entries for this Health Region corresponded to White/Caucasian individuals, who reported an overall 83% vaccination rate (Supplementary Table A-6). However, the interaction effect of Health Region and race was also significant in the case of individuals identifying as South Asian or other minorities not included in the survey [@tbl-model]. In this case, the results of the interaction term in the model indicate that the odds of vaccination for those within the South Asian and Other minority groups in the West Region decreased when compared to the other Health Regions (Supplementary Figure A-4).

According to Ontario Health, 13.2% of the population in the West Health Region identifies as a visible minority, whereas 2.5% identifies as Indigenous [@ontariohealth]. In the case of this analysis, the estimated lower odds are likely to be explained from a socio-economic perspective. In fact, 50% of the answers from this region in the survey came from the former LHINs of Hamilton Niagara Haldimand Brant, and Erie St. Clair, both which are among the regions of Ontario with the highest proportion of their population (more than 20%) in the lowest income quintile [@buajitti2018] (Supplementary Table A-7). Therefore, this result partly reinforces the well-known existing association between low vaccination rates and income, but it additionally indicates that there were intra-regional differences in vaccination. Interestingly, a disproportionate number of COVID-19 cases and low vaccination rate (under 50%) have been previously reported in the South Asian community of Ontario [@anand2022]; in this regard, our result provides additional context  by showing that within the South Asian community, there were differences in vaccination uptake across Ontario. Moreover, because significant lower odds of vaccination were also identified  other minority groups, this provides a rationale for future studies that explore how vaccination uptake varies across different minority groups within Ontario and other Canadian provinces.

There are some limitations to the present study. First, the data collection design, which allowed respondents to withdraw from the survey at any point, resulted in a high number of unique entries in the survey with multiple missing answers. Because we focused on entries that had complete observations in the covariates of interest for our analysis, it is possible that some information was not considered by excluding observations that had information in other variables (such as work from home, or number of persons in the household). However, we attempted to minimize this possibility by correcting the dataset using information from the Census. More granular corrections, which for example could be based on demographic information by municipality, could be used in the future to obtain a more accurate approximation to the population totals of the province. Additionally, the results in this study are based on self-reported data, where the risk of bias exist. Despite this, because in the context of COVID-19 it has been shown that good agreement exists between self-reported and documented vaccination status [@stephenson2022], the effect of self-reported bias is likely to not be significant in this case.

Finally, it is likely that there have been differences in vaccination across the province as more doses of the vaccine were administered and as successive variants emerged. Because this study focused only on vaccination status regarding the first dose of the vaccine within a relatively short time window, it can only provide a snapshot of the societal dynamics behind the pandemic. Nonetheless, the results presented here can serve as a starting point to motivate future longitudinal research that aims to quantify geographical differences within vulnerable segments of the population, and that can be used to inform the development of public health policies within the province of Ontario or across other provinces that aim to minimize disparities in health access.

### 2023-03-02 manuscript/appendix.qmd
Abstract, discussion, conclusion

```{r,interaction-race-income}
#| echo: false
#| warning: false
#| fig-height: 4
#| fig-width: 6
#| fig-cap: Interaction plot between Race and Income

@@

```{r,interaction-race-hr}
#| echo: false
#| warning: false
#| fig-height: 4
#| fig-width: 6
#| fig-cap: Interaction plot between Race and Health Region

@@

clean_data %>% subset (LHIN=="Hamilton Niagara Haldimand Brant (Hnhb)")%>%
tbl_summary(include=c(Race),
            sort = all_categorical() ~ "frequency") %>% 
  bold_labels()

>>>>

```{r,interaction-race-income}
#| echo: false
#| warning: false
#| fig-height: 4
#| fig-width: 6
#| fig-cap: Estimated marginal means for the combination between Race and Income in the regression model.

@@

```{r,interaction-race-hr}
#| echo: false
#| warning: false
#| fig-height: 4
#| fig-width: 6
#| fig-cap: Estimated marginal means for the combination between Race and Health Region in the regression model.

@@

# clean_data %>% subset (LHIN=="Hamilton Niagara Haldimand Brant (Hnhb)")%>%
# tbl_summary(include=c(Race),
#             sort = all_categorical() ~ "frequency") %>% 
#   bold_labels()
# 
# 
# clean_data %>% subset (LHIN=="Hamilton Niagara Haldimand Brant (Hnhb)")%>%
# tbl_summary(by=first_dose,
#             percent="row",
#   include=c(Race),
#             sort = all_categorical() ~ "frequency") %>% 
#   bold_labels()
# 
# clean_data %>% subset (LHIN=="Erie St. Clair")%>%
# tbl_summary(by=first_dose,
#             percent="row",
#   include=c(Race),
#             sort = all_categorical() ~ "frequency") %>% 
#   bold_labels()
```

### 2023-03-02 manuscript/main.qmd
Abstract, discussion, conclusion

@@

@@

>>>>

The COVID-19 pandemic continues to be a worldwide public health concern. Although vaccines against this disease were rapidly developed, vaccination uptake has not ben equal across all the segments of the population. In particular, it has been shown that there have been differences in vaccine uptake across different segments of the population. However, there are also differences in vaccination across geographical areas, which might be important to consider in the development of future public health policies against COVID-19. In this study, we examined the relationship between vaccination status (having received the first dose of a COVID-19 vaccine), and different socio-economic and geographical factors. Our results show differences in vaccination due to race/ethnicity, income, Health Regions (geographical areas used for health service access in Ontario), and their interactions. In particular, we show
that individuals who identified as Arab/Middle Eastern, Black, or Latin American, had significantly lower odds of vaccination than White/Caucasian individuals (ORs=0.31, 0.32, 0.28, and _p_=0.004,_p_<0.001 and _p_=0.004, respectively), and that individuals with a household income below CAD 25,000 who identified as Arab/Middle Eastern (OR=3.05, _p_=0.013), Black (OR=3.19, _p_=0.004), Latin American (OR=2.80, _p_=0.041), or that belonged to other minority groups  (OR=4.59, _p_<0.001) had higher odds of vaccination than individuals from the same racial/ethnic group in higher income brackets. Finally, we also identified lower odds of vaccination within different minority groups in West Health Region. This study shows that there is an ongoing need to better understand and address differences in vaccination uptake across diverse segments of the population that have been largely impacted by the pandemic.

@@

Covid-19, vaccination, survey, socio-economic factors, visible minorities.

@@

```{r}
#| child: 05-conclusion.qmd
```
